---
title: "K-Means Clustering Algorithm Practice"
output: html_notebook
---

## Part 1

Create two groups of clusters, the answers to the questions and regions where people grew up. 

```{r}
library(dplyr)
library(tidyr)  
library(ggplot2)
```

```{r}
df <- read.csv("HUDK4050_2017_Cluster_Survey.csv")
```

```{r}
df_answers <- select(df, -first, -last, -city.grew, -state.grew, -country.grew)
```

```{r}
df_answers$cat <- as.integer(df_answers$cat)
df_answers$gif.jiff <- as.integer(df_answers$gif.jiff)
```

```{r}
fit <- kmeans(df_answers, 2)
df_cluster2 <- data.frame(df_answers, fit$cluster)
fit <- kmeans(df_answers, 3)
df_cluster3 <- data.frame(df_answers, fit$cluster)
fit <- kmeans(df_answers, 4)
df_cluster4 <- data.frame(df_answers, fit$cluster)
fit <- kmeans(df_answers, 5)
df_cluster5 <- data.frame(df_answers, fit$cluster)
fit <- kmeans(df_answers, 6)
df_cluster6 <- data.frame(df_answers, fit$cluster)
fit <- kmeans(df_answers, 7)
df_cluster7 <- data.frame(df_answers, fit$cluster)
fit <- kmeans(df_answers, 8)
df_cluster8 <- data.frame(df_answers, fit$cluster)
fit <- kmeans(df_answers, 9)
df_cluster9 <- data.frame(df_answers, fit$cluster)
fit <- kmeans(df_answers, 10)
df_cluster10 <- data.frame(df_answers, fit$cluster)
```

```{r}
# Create an elbow plot to determine what centers for kmeans
# https://en.wikipedia.org/wiki/Determining_the_number_of_clusters_in_a_data_set
y <- c(var(df_cluster1$fit.cluster), var(df_cluster2$fit.cluster), var(df_cluster3$fit.cluster), var(df_cluster4$fit.cluster), var(df_cluster5$fit.cluster), var(df_cluster6$fit.cluster), var(df_cluster7$fit.cluster), var(df_cluster8$fit.cluster), var(df_cluster9$fit.cluster), var(df_cluster10$fit.cluster))
x <- c(1:10)
plot(x,y, type = "p", main = "Variance vs. K-Means Center", sub = "Elbow Method to Determine K-Means Center", xlab = "Clustering Size", ylab = "Variance")

```

```{r}
df_region <- select(df, country.grew)
count(df_region, country.grew)
```
```{r}
df_region$country.grew[df_region$country.grew == 'USA'] <- 'United States'
df_region$country.grew[df_region$country.grew == 'USA '] <- 'United States'
count(df_region, country.grew)
```

```{r}
plot <- select(df_cluster6, fit.cluster)
plot <- bind_cols(plot, df_region)
plot$fit.cluster <- as.numeric(plot$fit.cluster)
plot$country.grew <- as.numeric(plot$country.grew)
names(plot) <- c("clustered_answers", "region")
plot <- arrange(plot, region)
```


##Part III

Create a visualization that shows the overlap between the two groups of clusters you created in part II.

visualize average totaled survey answers by region, by cluster

```{r}
print(plot)
```

```{r}
ggplot(clustered_answers, aes(region, clustered_answers, colour = region)) + geom_count() + xlab("Region") + ylab("Survey Answer Clusters")
```

Summary of what I did:
1. df_cluster6$fit.cluster clustered the survey answers into a k-means center of 6. I chose 6 as the result of the elbow method.
2. manually categorized the countries as regions.
Region Key:
-- China	= 1 (n=16)
-- Indonesia	= 2 (n=1)			
-- Taiwan = 3 (n=1)			
-- United States = 4 (n=7)	
3. bind df_cluster6$fit.cluster and as.numeric(regions)
4. ggplot + geom_count() gave me a graph where:
x-axis: region
y-axis: survey answers
points: are small or big depending on n=

To interpret:
If you look at the data by region, regions 1 (China) and 4 (US) have the most populated data.
If you compare regions 1 and 4, region 1 has two distinct clusters (1 and 4 on the y axis) and region 4 has two distinct clusters (1 and 2).
Regions 1 and 4 share cluster 1 as one of their two distinct clusters.
All regions share cluster 1 as a cluster.
Regions 1 and 4 have two overlapping clusters (1 and 4) but do not overlap with clusters 3, 4, 5 and 6.
Clusters 5 and 6 for region 4 have small n values.
Clusters 2 and 3 for region 1 have small n values.

Recommendation:
I would recommend looking for patterns in the clustered survey answers determined by R for further interpretation as it might pertain to region differentiation.


```{r}
count(df_cluster6, fit.cluster)
```

